### 2019-08-28 README.Rmd
Cleaned up README and vignette
spell/grammer check README and vingette

In 2012 KEGG released its first application programming interface (API), and has been
adding features and functionality ever since. There are benefits to using an API.
First, API's like KEGG's allow users to perform customized analyses with the most
up-to-date versions of the data contained in the data base. In addition, accessing the
KEGG API is very easy using statistical programming tools like R or Python and
integrating data pulls into user's code makes the program reproducible. To 
further enforce reproducibilty *diffEnrich* adds a date and KEGG realease tag to
all data files it generates from accessing the API. For update histories and release 
notes please go [here](https://www.kegg.jp/kegg/rest/).

@@

A recent [study](https://onlinelibrary.wiley.com/doi/full/10.1111/acer.13766) explored 
the hepatic mechanism for removal of toxic lipid aldehydes via conjugation with glutathione (Cheng et al., [2001](https://www.ncbi.nlm.nih.gov/pubmed/11488593?dopt=Abstract);
Gallagher et al., [2007](https://www.ncbi.nlm.nih.gov/pubmed/17553661?dopt=Abstract)) 
and the primary enzyme that catalyzes this conjugation, glutathione S‐transferase A4‐4 (GSTA4).
Specifically, the researchers examin the role of the GSTA4 gene on protein 
carbonylation and the progression of liver injury in a model consisting of long‐term 
(116 days) chronic ethanol (EtOH) consumption followed by a single EtOH binge.

@@

### Step 1: Collect and clean pathways from KEGG API
First we will use the *get_kegg* function to access the KEGG REST API and download the data sets required to perform 
our downstream analysis. This function takes two arguments. The first, 'species' is required and is the species of 
interest. Currently, *diffEnrich* supports three species, and the argument is a character string using the KEGG
code [REF](https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf): 
Homo sapiens (human), use 'hsa'; Mus musculus (mouse), use 'mmu', and Rattus norvegicus (rat), use 'rno'. The second,
'path' is also passed as a character string, and is the path of the directory in which the user would like to
write the data sets downloaded from the KEGG REST API. If the user does not provide a path, the data sets will 
be automatically written to the current working directory using the *here::here()* functionality. These data 
sets will be tab delimited files with a name describing the data, and for reproducibility, the date they were generated
and the version of KEGG when the API was accessed. In addition to these flat files, *get_kegg* will also
create a named list with the three relevant KEGG data sets. The names of this list will describe the the data set.
For a ddetailed descritpion of list elements use *?get_kegg*.

@@

In this step we will use the *pathEnrich* function to identify KEGG pathways that are enriched (i.e. over-represented) based on two gene lists of interest. The ENTREZ gene IDs. User gene lists must also be charater vectors and be formatted as ENTREZ gene IDs. The *clusterProfiler* offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.github.io/clusterProfiler-book/chapter5.html#supported-organisms).

@@

*plotFoldEnrichment* generates a grouped bar plot using ggplot2 and the
*ggnewscale* package. KEGG pathways are plotted on the y-axis and fold
enrichment is plotted on the x-axis. each KEGG pathway has a bar plotting
its fold enrichment in list 1 (red) and its fold enrichment in list 2 (blue).
The transparency of the bars correspond to the adjusted p-value (FDR) for the
pathway's enrichment in the given list. The p-value presented as text to the
right of each pair of bars is the adjusted p-value (user defined: default is FDR) associated with the
differential enrichment of the pathway between the two lists, and the pathways
are ordered from top to bottom by this p-value (i.e. smallest p-value on top
of plot, and largest p-value on bottom of plot). Finally, the number of genes used
for analysis from each gene list (recall that this number may not be the same as the number of
genes in the user's original list) are reported below their repsective p-values
in the legend.

@@

## References

>>>>

In 2012 KEGG released its first application programming interface (API), and has been
adding features and functionality ever since. There are benefits to using an API.
First, API's like KEGG's allow users to perform customized analyses with the most
up-to-date versions of the data contained in the data base. In addition, accessing the
KEGG API is very easy using statistical programming tools like R or Python and
integrating data pulls into user's code makes the program reproducible. To 
further enforce reproducibilty *diffEnrich* adds a date and KEGG release tag to
all data files it generates from accessing the API. For update histories and release 
notes please go [here](https://www.kegg.jp/kegg/rest/).

@@

A recent [study](https://onlinelibrary.wiley.com/doi/full/10.1111/acer.13766) explored 
the hepatic mechanism for removal of toxic lipid aldehydes via conjugation with glutathione (Cheng et al., [2001](https://www.ncbi.nlm.nih.gov/pubmed/11488593?dopt=Abstract);
Gallagher et al., [2007](https://www.ncbi.nlm.nih.gov/pubmed/17553661?dopt=Abstract)) 
and the primary enzyme that catalyzes this conjugation, glutathione S‐transferase A4‐4 (GSTA4).
Specifically, the researchers examine the role of the GSTA4 gene on protein 
carbonylation and the progression of liver injury in a model consisting of long‐term 
(116 days) chronic ethanol (EtOH) consumption followed by a single EtOH binge.

@@

### Step 1: Collect and clean pathways from KEGG API
First we will use the *get_kegg* function to access the KEGG REST API and download the data sets required to perform 
our downstream analysis. This function takes two arguments. The first, 'species' is required and is the species of 
interest. Currently, *diffEnrich* supports three species, and the argument is a character string using the KEGG
code [REF](https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf): 
Homo sapiens (human), use 'hsa'; Mus musculus (mouse), use 'mmu', and Rattus norvegicus (rat), use 'rno'. The second,
'path' is also passed as a character string, and is the path of the directory in which the user would like to
write the data sets downloaded from the KEGG REST API. If the user does not provide a path, the data sets will 
be automatically written to the current working directory using the *here::here()* functionality. These data 
sets will be tab delimited files with a name describing the data, and for reproducibility, the date they were generated
and the version of KEGG when the API was accessed. In addition to these flat files, *get_kegg* will also
create a named list with the three relevant KEGG data sets. The names of this list will describe the the data set.
For a detailed description of list elements use *?get_kegg*.

@@

In this step we will use the *pathEnrich* function to identify KEGG pathways that are enriched (i.e. over-represented) based on two gene lists of interest. The ENTREZ gene IDs. User gene lists must also be character vectors and be formatted as ENTREZ gene IDs. The *clusterProfiler* offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.github.io/clusterProfiler-book/chapter5.html#supported-organisms).

@@

*plotFoldEnrichment* generates a grouped bar plot using ggplot2 and the
*ggnewscale* package. KEGG pathways are plotted on the y-axis and fold
enrichment is plotted on the x-axis. each KEGG pathway has a bar plotting
its fold enrichment in list 1 (red) and its fold enrichment in list 2 (blue).
The transparency of the bars correspond to the adjusted p-value (FDR) for the
pathway's enrichment in the given list. The p-value presented as text to the
right of each pair of bars is the adjusted p-value (user defined: default is FDR) associated with the
differential enrichment of the pathway between the two lists, and the pathways
are ordered from top to bottom by this p-value (i.e. smallest p-value on top
of plot, and largest p-value on bottom of plot). Finally, the number of genes used
for analysis from each gene list (recall that this number may not be the same as the number of
genes in the user's original list) are reported below their respective p-values
in the legend.

@@
